preference

### 2020-09-09 analyses/FL_abund_GAM.Rmd
update

** One thing I need to go back and reconsider is that right now, I think forbidden links (i.e. links between species that never co-occurred) are given a zero for preference. A better approach would be to set their preference to a neutral value, which would be 1 in my case, since exp(0) = 1 and my preference values are calculated as exp(residual) of my linear model of visitation ~ abundance. **

@@

```{r}
### Global model
weighted_fl_abundance.global <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(year, bs = "re") +
                                      s(site, bs = "re") +
                                        te(yday, elev.mean, 
                                         by = year,
                                                id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2) +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=1),
                                    data = fl_abund_wt,
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

@@

weighted_fl_abundance.global3 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(year, bs = "re") +
                                      s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                              id = 0,
                                              bs= c("gp", "tp"), 
                                              k=c(10, 10), 
                                              m=2) +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=1),
                                    data = fl_abund_wt,
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

weighted_fl_abundance.global4 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(year, bs = "re") +
                                      s(site, bs = "re") +
                                      te(yday, elev.mean, year,
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp", "re"), 
                                                k=c(5, 5), 
                                                m=1),
                                    data = fl_abund_wt,
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

AIC(weighted_fl_abundance.global, weighted_fl_abundance.global2, weighted_fl_abundance.global3, weighted_fl_abundance.global4)

@@

weighted_fl_abundance.global.sum <- summary(weighted_fl_abundance.global)
weighted_fl_abundance.global.sum2 <- summary(weighted_fl_abundance.global2)
weighted_fl_abundance.global.sum3 <- summary(weighted_fl_abundance.global3)
weighted_fl_abundance.global.sum3 <- summary(weighted_fl_abundance.global4)

print(plot(weighted_fl_abundance.global, allTerms = TRUE), pages = 4)
print(plot(weighted_fl_abundance.global2, allTerms = TRUE), pages = 4)
print(plot(weighted_fl_abundance.global3, allTerms = TRUE), pages = 4)
print(plot(weighted_fl_abundance.global4, allTerms = TRUE), pages = 4)

@@

### 2010
weighted_fl_abundance.2010 <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + 
                                    management + 
                                    #slope.est +
                                    #s(site, bs = "re") +
                                    te(yday, elev.mean, 
                                       by = bb.sp,
                                       bs= c("gp", "tp"), 
                                       k=c(10, 10), m=2),
                                  data = filter(fl_abund_wt, year == 2010),
                                  discrete = TRUE,
                                  family = "tw",
                                  select = TRUE,
                                  method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2010)
weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 1)
weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2010.ck + l_gridCheck2D(gridFun = mean)

### 2011
weighted_fl_abundance.2011 <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + 
                                    #s(transect, bs = "re") +
                                    management + 
                                    #slope.est +
                                    #s(site, bs = "re") +
                                    te(yday, elev.mean, 
                                       by = bb.sp,
                                       bs= c("gp", "tp"), 
                                       k=c(10, 10), m=2),
                                  data = filter(fl_abund_wt, year == 2011),
                                  discrete = TRUE,
                                  family = "tw",
                                  select = TRUE,
                                  method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2011)
weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 1)
weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2011.ck + l_gridCheck2D(gridFun = mean)

### 2012
weighted_fl_abundance.2012 <- bam(fl.abund.wt ~ s(bb.sp, bs = "re") + 
                                    #s(transect, bs = "re") +
                                    management + 
                                    #slope.est +
                                    s(site, bs = "re") +
                                    te(yday, elev.mean, 
                                       by = bb.sp,
                                       bs= c("gp", "tp"), 
                                       k=c(10, 10), m=1),
                                  data = filter(fl_abund_wt, year == 2012),
                                  discrete = TRUE,
                                  family = "tw",
                                  select = TRUE,
                                  method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2012)
weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 1)
weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2012.ck + l_gridCheck2D(gridFun = mean)
```

>>>>

### Would these models be better specified using gamm4? Would that let me handle the crossed and nested random effects better?

@@

```{r}
### A gamm4 version
weighted_fl_abundance.gamm4 <- gamm4(log(fl.abund.wt + 0.001) ~
                                       t2(yday, elev.mean, 
                                          by = bb.sp,
                                          #id = 0,
                                          bs= c("gp", "tp"), 
                                          k=c(8, 8), 
                                          m=1),
                                     random = ~(1 | site:year),
                                     data = fl_abund_wt,
                                     discrete = TRUE,
                                     family = "gaussian",
                                     select = TRUE,
                                     method = "fREML") %>% getViz(nsim = 500)

### Global model
#### My concern with using a global model is that, because I have so much higher sampling density in 2012, the global model will basically be just a 2012 model with 2010 and 2011 in the residuals, so to speak. I think this is at least partly mitigated by included a global by = year smooth, but I still don't quite understand how hierarchical effects relate to one another in a GAM. I'm also concerned that if I include the by = year smooth, it will be too concurve with my by = bb.sp smooths, and it will be hard to interpret the pattern of resource availability per species.
weighted_fl_abundance.global <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(year, bs = "re") +
                                      s(site, bs = "re") +
                                        te(yday, elev.mean, 
                                         by = year,
                                                id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(5, 5), 
                                                m=2) +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=1),
                                    data = fl_abund_wt,
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

@@

AIC(weighted_fl_abundance.global, weighted_fl_abundance.global2)

@@

weighted_fl_abundance.global.sum <- summary(weighted_fl_abundance.global)
weighted_fl_abundance.global.sum2 <- summary(weighted_fl_abundance.global2)

print(plot(weighted_fl_abundance.global, allTerms = TRUE), pages = 4)
print(plot(weighted_fl_abundance.global2, allTerms = TRUE), pages = 4)

@@

weighted_fl_abundance.global2.ck <- check2D(weighted_fl_abundance.global2, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.global2.ck + l_gridCheck2D(gridFun = mean)

### 2010
weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2010)
weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2010.ck + l_gridCheck2D(gridFun = mean)

### 2011
weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2011),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2011)
weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2011.ck + l_gridCheck2D(gridFun = mean)

### 2012
weighted_fl_abundance.2012 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2012),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2012)
weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2012.ck + l_gridCheck2D(gridFun = mean)
```
